Vestibular Contributions to Spatial Perspective Transformations: From Sensory Inference To Imagined Self-Motion

### 2017-01-27 manuscript.Rmd
draft introduction

abstract: >
        Recent research has indicated that the vestibular network is involved in various cognitive tasks involving mental spatial transformations. However, the nature of this involvement remains unclear.  Here, we provide a computational framework for vestibular cognition and discuss this in the context of perspective transformations, which require a simulated rotation of the self. We explore how the ability to model the effects of body rotation using vestibular input might enable the ability to imagine body rotations. We approach this problem by extending previous models of vestibular sensory processing to include the ability to model the sensory consequences of self-initiated movements, and discuss how this can be further extended to allow mentally simulated rotations of the self. We ... Predicting the sensory input is key to understanding imagined rotations. Mental imagery goes beyond prediction, rather, it involves the use of counterfactual queries in a probabilistic model.

@@

A number of recent studies have demonstrated that the vestibular system, which deals with sensory signal relating to movements of the head, is involved in such spatial transformations [@Lenggenhager:2008et, @Deroualle:2015gk, @Falconer:2012cl, @Gardner:2016kd]. This is demonstrated in the form of interference effects between mental transformations and concurrent sensory stimulation. However, to date, there has little attempt to understand the involvement of the vestibular system in higher-level cognitive abilities from a computational perspective. In this paper, we focus specifically on the role of the vestibular system for spatial perspective transformations, and we draw the link between probabilistic computations performed by the vestibular system in the service of real-time interaction with the world, and the brain's ability to run mental simulations for the purpose of adopting the perspective of another person.

Consider the use of deictic pronouns, such as "this" or "that" and "left" or "right". These terms only assume a their meaning when placed within a certain frame of reference (FOR). For example, in a dialogue between two speakers, the statement "I would like that one", or "I would like the one on the right" can only be understood under the assumption that the deictic terms "that" and "right" are used in an egocentric FOR. In addition, both interlocutors must be able to access the other's FOR with respect to an external FOR. In other words: In order to give instructions, the speaker must be able to mentally assume the addressee's position in space. A further example that requires a similar spatial transformation is when two people want to perform a coordinated action, such as lifting a heavy object. This requires having knowledge of the other's potential actions in order to adjust one's own actions, and this again requires knowledge of the other's position with an external FOR. A judgement about what kind of actions the other person can perform requires an embodied simulation using a representation of one's own body schema [@Pezzulo:2013iy]. Giving an instruction to another person thus requires running an embodied simulation in order to infer what kind of motor commands the other person should perform __[Blakemore and Decety, 2001; Wolpert et al., 2003; Jeannerod, 2006; Pezzulo et al., 2007, 2013; Dindo et al., 2011; Pezzulo, 2011a,b]__. This embodied simulation is performed using one's own body schema. Communicating this information to the other actor requires knowing the position of the addressee's FOR relative to the external reference frame, and the difference between the speaker and the addressee's egocentric reference frames. Perspective taking thus can be conceived of as a aligning a representation of one's body with that of another person in order to access spatial information from a viewpoint other than one's own egocentric viewpoint. In order to achieve this alignment, one must perform translations and rotations of one's own egocentric reference frame. Processing translations and rotations of the head and body is the core domain of the vestibular system, and thus is of particular interest for research into mental spatial transformations. 
Spatial transformations in 6 dimensions (translation along 3 axes, and rotations around 3 axes) are highly complex, and non-linear, and we propose that the brain adopts the strategy of simulating motion of the body through space in order achieve a desired target position. This effectively amounts to a running a simulation of dynamical system, consisting of simulated limb movements and the resulting sensory consequences. Figure \@ref(fig:spt-schematic) (a) shows a simple scenario in which a speaker wants to give an addressee, standing opposite, verbal instructions about an action to perform on an external object. The speaker must determine the most suitable action, given the addressee's position relative to the external object, and in order to achieve this goal, the speaker must align himself with the addressee, so that he can use his own body-centred reference frame to quickly simulate the best action. In this example, alignment with the addressee's reference frame consists of a forwards translation and a rotation around the vertical axis. This can be described as an imagined movement, in which the goal is achieve a specific position.

@@

In order to do this, the speaker must first determine what type of movement to perform, and the direction, velocity and duration that lead to the desired outcome. This means that in order to know what to simulate, the speaker must first work backwards from the desired outcome, and infer the actions that are most likely to lead to that outcome. This is known in the field of motor control as the inverse problem [@Wolpert:1998td]. In order to achieve the desired position, the motor commands can then be simulated, and the expected sensory consequences can be estimated; this is known as the forward problem.  Figure \@ref(fig:spt-schematic) (b) shows that if the goal is to simply simulate the perceptual experience of rotation around the body's vertical (yaw) axis, rather than project a representation of one's body onto a target, the problem is simplified; the actor can perform forward inference without the need to determine the optimal motor action. As demonstrated by @Penny:2013iv, these computations can be performed in a common probabilistic model. Interestingly, this has been discussed in the context of reinforcement learning, and planning in humans has been described as reverse inference [@Botvinick:2012fe]. Further, [@Pickering:2014bo] discuss the relationship between control-theoretic notions of inverse/forward models, known collectively as internal models, and and more general concepts of a generative model in a hierarchical Bayesian framework.

Performing a simulation of body movements and estimating the sensory consquences amounts to simulating the behaviour of a dynamical system using a forward model; this requires a generative model of the system.The aim of this paper is to specifically investigate the generative model that may be used for mental spatial transformations. The ability to run embodied simulations is often claimed to be grounded in the relevant sensorimotor circuits, and we claim that the vestibular system is ideally suited to investigate this connection, because the generative models used by the vestibular system for dynamic sensory inference have been extensively investigated using approaches from optimal control theory, and more recently, dynamic Baysian inference. The vestibular system faces a number of challenges in interpreting the sensory measurements provided by the semi-circular canals (SCC) and otolith organs. On the one hand, these signals  are very noisy, and ambiguous [@PaulRMacNeilage:2008bv]; the otoliths respond both to translations and tilt of the head relative to gravity, and in order to disambiguate between translation and tilt, the brain must perform sophisticated computations at the level of the brain stem. It has become clear that the nature of these computations is probabilistic, and that the brain makes use of strong prior beliefs in order to interpret these sensory signals. A further challenge faced by the vestibular system is that it needs to distinguish between sensory signals that are the result of active motion, and those resulting from external forces [@Cullen:2009dc, @Cullen:2014gx]. The ability to distinguishing re-afferent from ex-afferent sensory signals is an essential capability of all organisms that move actively [@Crapse:2008jc].

Furthermore, the apparent connection between mental imagery and predictive processing has repeatedly been alluded to [@Moulton:2009ij, @Clark:2012vf, @Gambi:2015gv], and @Grush:2004wx claims that mental imagery is performed by emulator circuits implementing a forward model. These ideas can also be addressed from a computational point of view in the context of dynamic Bayesian inference, where there are multiple roles for sensory predictions, over different time scales. This was recently discussed by [@Tian:2012ui] in the context of imagined speech.

The putative link between predictive processing and higher cognitive processing has a firm basis in predictive processing accounts of sensory inference [Friston]. The predictive processing framework has been applied to almost all sensory modalities [CITATIONS: pain, proprioception, auditory , visual, etc.], and the picture that has emerged in recent years is that the brain employs sophisticated inference algorithms in order to "interpret" the stream of incoming sensory data. The brain seems to be actively engaged in formulating hypotheses, and testing these on sensory data; this has been extended to the concept of active inference [Friston], in which an agent attempts to either change it's internal model of world, or to interven on the world in order to fulfil sensory predictions. We claim that the vestibular modality is particularly well-suited to investigating predictive processing at multiple hierarchical scales, from the lowest levels of sensory inference to higher levels, where goals, intentions and beliefs of the organism become relevant. A case in point is the study by @Wertheim:2001cb, who provided evidence that participants' perceptions of tilt induced by vestibular stimulation are shaped to a large extent by the beliefs about possible motions they formed prior to taking part in the experiement.

Interference effects between higher-level cognitive processes, such as SPT and mental imagery, and sensory inference, are thought to occur as a result of shared circuitry [@Deroualle:2014ho, @Nigmatullina:2015gh]. @Ellis:2017fg recently claimed that these types of interference effects might arise as a result of cognitive processing exerting an influence on circuits performing sensory inference via shared higher-level priors. In this paper, we pursue this claim, and we attempt to the grounding of cognitive abilities in the computations performed by low-level sensory circuits in the service of dynamic real-time interactions with the world. A problem that needs to addressed is whether models that are applicable to lower-level sensory inference are scalable, in the sense that they are able to represent hierarchically structured knowledge, and can be applied to tasks requiring compositionality (?) [@Kwisthout:2016fx, @Griffiths:2012ii]. In the field of cognitive science, sophisticated model of structured representations using Bayesian networks [@Pearl:1988wz] have been developed [@Tenenbaum:2011hd, @Lucas:2010tu], with applications to concept learning [CITATION], inductive biases in reasoning, in cognitive models of development [@Perfors:2011ex], language [@Goodman:2016go] and social cognition [@Baker:2009kx].
@Kwisthout:2016fx make the point that predictive processing accounts are situated at low levels in the processing hierarchy, usually at the level of sensory inference or at the level describing kinematics. In this paper, we address this issue by starting from a model developed for low-level sensory processing, incorporating basic abilities to predict sensory consequences of movement based on knowledge about the kinematics, and extending this to account for interactions between thought processes, operationalized as mental simulations of sensorimotor experience, and sensory processing.

> TODO: CITE [@Battaglia:2013fm]

To our knowledge, few studies have addressed the idea that higher level cognitive processes are grounded in the circuits performing sensory inference. We claim that if cognition is grounded in sensory inference, and that this is to be experimentally studied by analyzing interactions between thought and perception, then it is vital that this be carried within the constraints of a theoretical (computational) framework. A suitable formalism for this is the framework of probabilistic graphical models, and within this framework, interactions between thought and perception can be modelled as effects of higher level priors and hyperpriors on inference at a computational level. Investigating inference at the level of algorithmic implementations might provide further insight.

@@

We investigate one specific type of head movement, a rotation of the head about the yaw axis, and discuss this in the context of the probabilistic computations that are performed when the goal is to:

@@

@@

## Previous work
Whereas early models of vestibular sensory processing were inspired by control theory [@Borah:1988fd, @Merfeld:1999cg], several recent studies have described vestibular sensory processing as dynamic Bayesian inference [@Laurens:2006be, @PaulRMacNeilage:2008bv, @Karmali:2012cv]. @Selva:2012gsa
provide a recent review of the relationship between optimal observer models and models that employ Kalman filtering, and show that these are largely equivalent. Both @Karmali:2012cv and @Laurens:2006be described vestibular sensory inference in terms of particle filtering, which is a particular algorithm that can be used to perform inference in state-space models. [@Doucet:2000bh, @Doucet:2009us]. Using their model, @Laurens:2006be were able to simulate perceptual responses to centrifugation and off-vertical axis rotations. A well-known phenomenon of vestibular perception is velocity storage, which is defined as the slower decay of perceptual and oculomotor responses, relative to the SCC afferent responses [@Raphan:1979hn], __[Robinson 1977]__. This is generally seen as evidence for the existence of an internal model in the processing of vestibular signals. @Karmali:2012cv focussed on one-dimensional angular velocity during yaw rotations; their model was able to explain the phenomenon of velocity storage as a consequence of particle filtering, with the time constant of velocity storage depending on the variability of the sensory afferent signals.

## Generative model for sensory inference
We proceed by providing a brief introduction to the concept of state-space models, and particle filtering, which is a particular algorithm that can be used to perform inference in state-space models. While Kalman filters provide the optimal solution when the restrictions of linear Gaussian state-space models are met, they are not easily extensible to non-linear models, which is relevant when considering multi-dimensional motion. Particle filtering can be applied to arbitrary nonlinear models. State-space models can be seen as special cases of more general Bayesian hierarchical models [@Bishop:2006ui; @Murphy:2012ua]. Following the same approach as @Penny:2013iv, we first reframe the generative model used for sensory inference as a dynamic probabilistic graphical model.
 <!-- from a statistical perspective, rather an engineering perspective -->
A state-space model is special case of a dynamic latent variable model, which is suitable for modelling the time-varying behaviour of a system in terms of latent, unobservable processes and noisy observations. As such, a state-space model can be viewed as a statistical model of the problem at hand.

For example, consider trying to estimate the velocity and position during a leftward head turn. The afferent SCC signals are considered to be proportional to the angular velocity of the head with a frequency band of approximately 0.1 – 10 Hz, which constitutes the range of natural head movements [@Grabherr:2008kk; @Carriot:2014fv]. The measurements are therefore observations of the true, unobservable angular velocity; however, they are contaminted by Gaussian noise, which is determined by characteristics of the sensory organs. Figure \@ref(fig:head-turn-data) shows a sequence of noisy measurements during a two-second turn of the head. The acceleration generating the movement is sinusoidal, given for each time step $t$ by $Asin(2\pi t/T)$, where $A$ is the amplitude of the movement, $T = 1/f$ is the duration of the movement, and $f$ is the frequency. This acceleration results in a peak velocity $\omega_{peak} = AT/\pi$ and a final angular position $\theta_{final} = AT^2/2\pi$.  Due to the inherent noisiness of the measurements, and the dynamic nature of the stimulus, the measurements need to be interpreted according to the context in which they occur.

@@

# motiondata_hidden <- motiondata %>%
#     gather(type, value = value, -time) %>%
#     filter(type != "observations") %>%
#     mutate(type = factor(type,
#                          levels = c("position",
#                                     "velocity",
#                                     "acceleration")))
# 
# p1 <- motiondata_hidden %>%
#         ggplot(aes(x = time, y = value)) +
#             geom_line(aes(linetype = type), size = 1.4) +
#             geom_point(data = motiondata,
#                        aes(x = time, y = observations),
#                        shape = 21, colour = "white",
#                        fill = "white", size = 6, stroke = 2) +
#             geom_point(data = motiondata,
#                        aes(x = time, y = observations),
#                        shape = 21, colour = "black",
#                        fill = "white", size = 4, stroke = 2) +
#             guides(linetype = guide_legend(keywidth = 2.5, keyheight = 1)) +
#             xlab("Time (sec)") +
#             ylab("Angular position (deg)") +
#             theme_apa() +
#             theme(legend.title=element_blank())
# p1
```

@@

# TODO
How do we explain higher-level anticipation/expectations in such a model? This form of anticipation is not the same as the 1-step ahead predictions made during inference, but rather implement higher level prior beliefs about the direction of movement, for example.

>>>>

abstract: >
        Mental spatial transformations are considered to be an essential cognitive ability. Recent research has indicated that the vestibular network is involved in such mental spatial transformations. However, the nature of this involvement remains unclear. We propose a computational framework for the involvement of the vestibular system in cognition and discuss this in the context of perspective transformations, which require a simulated rotation of the self. We explore how the ability to anticipate the sensory consequences of self-initiated actions might underlie the ability to imagine body rotations. We approach this problem by extending previous models of vestibular sensory inference to include the ability to model the sensory consequences of self-initiated movements, and discuss how this can be further extended to allow mentally simulated rotations of the self. Our model is able to explain various source of interference between simulations of self-rotation and sensory inference.

@@

<!-- [ok] -->
A number of recent studies have demonstrated that the vestibular system, which processes sensory signal relating to angular and linear movements of the head, is involved in such spatial transformations [@Lenggenhager:2008et, @Deroualle:2015gk, @Falconer:2012cl, @Gardner:2016kd]. This is demonstrated in the form of interference effects between mental transformations and concurrent sensory stimulation. To date, however, there has little attempt to understand the involvement of the vestibular system in higher-level cognitive abilities in terms of the computational mechanisms.
<!-- [ok] -->
Consider the use of deictic pronouns, such as "this" or "that" and "left" or "right". These terms only assume a meaning relative to a certain frame of reference (FOR). For example, in a dialogue between two speakers, the statement "I would like that one", or "I would like the one on the right" can only be understood if both interlocutors are able to access the other's FOR with respect to an external FOR. A further example that requires a similar spatial transformation is when two people want to perform a coordinated action, such as lifting a heavy object. This requires having knowledge of the other's potential actions in order to adjust one's own actions, and this again requires knowledge of the other's position within an external FOR. A judgement about what kind of actions the other person can perform requires an embodied simulation using a representation of one's own body schema [@Pezzulo:2013iy] and giving an instruction to another person thus requires running an embodied simulation in order to infer what kind of motor commands the other person should perform __[Blakemore and Decety, 2001; Wolpert et al., 2003; Jeannerod, 2006; Pezzulo et al., 2007, 2013; Dindo et al., 2011; Pezzulo, 2011a,b]__. But this again requires information about the addressee's FOR relative to the external reference frame, and the difference between the speaker and the addressee's egocentric reference frames. Perspective taking can be conceived of as a aligning a representation of one's body with that of another person in order to access spatial information from a viewpoint other than one's own egocentric viewpoint. In order to achieve this alignment, one must perform translations and rotations of one's own egocentric reference frame. Processing translations and rotations of the head and body is the core domain of the vestibular system, and thus is of particular interest for research into mental spatial transformations. 
<!-- [ok] -->
Spatial transformations in 6 dimensions (translation along 3 axes, and rotations around 3 axes) are highly complex, and non-linear, and it has been proposed that the brain adopts the strategy of simulating motion of the body through space in order achieve a desired target position [CITATION]. This effectively amounts to a running a simulation of dynamical system, consisting of simulated limb movements and the resulting sensory consequences. Figure \@ref(fig:spt-schematic) (a) shows a simple scenario in which a speaker wants to tell an addressee, standing opposite, to pick something up from the floor. The speaker must determine the most suitable action by embodied simulation, given the addressee's position relative to the external object, and in order to achieve this goal, the speaker must align himself with the addressee, so that he can use his own body-centred FOR. In this example, alignment with the addressee's reference frame consists of a forwards translation and a rotation around the vertical axis. This can be described as an imagined movement, in which the goal is achieve a specific position.

@@

In order to do this, the speaker must first determine what type of movement to perform. This means that in order to know what to simulate, the speaker must first work backwards from the desired outcome, and infer the actions that are most likely to lead to that outcome [@Hamrick:2014wq]. In the field of motor control, this is known as the inverse problem [@Wolpert:1998td]. In order to achieve the desired position, the motor commands can then be simulated, and the expected sensory consequences can be estimated; this is known as the forward problem. 
Figure \@ref(fig:spt-schematic) (b) shows that if the goal is to deliberatively simulate the perceptual experience of rotation around the body's vertical (yaw) axis, rather than project a representation of one's body onto a target, the problem is simplified; the actor can perform forward inference without the need to determine the optimal motor action. 
As demonstrated by @Penny:2013iv, these computations can be performed in a common probabilistic model. Interestingly, this has been discussed in the context of reinforcement learning, and planning in humans has been described as reverse inference [@Botvinick:2012fe]. Further, [@Pickering:2014bo] discuss the relationship between control-theoretic notions of inverse/forward models, known collectively as internal models, and and more general concepts of a generative model in a hierarchical Bayesian framework.
<!-- [ok] -->
Simulating the behaviour of a dynamical system using a forward model requires a generative model of the system.

The aim of this paper is to specifically investigate the generative model that may be used for mental spatial transformations. The ability to run embodied simulations is often claimed to be grounded in the relevant sensorimotor circuits, and we claim that the vestibular system is ideally suited to investigate this connection, because the generative models used by the vestibular system for dynamic sensory inference have been extensively investigated using approaches from optimal control theory, and more recently, dynamic Baysian inference [Oman, CITATIONS, @PaulRMacNeilage:2008bv].

The vestibular system faces a number of challenges in interpreting the sensory measurements provided by the semi-circular canals (SCC) and otolith organs. On the one hand, these signals  are very noisy, and ambiguous; the otoliths respond both to translations and tilt of the head relative to gravity, and in order to disambiguate between translation and tilt, the brain must perform sophisticated computations. The emerging picture is that the nature of these computations is probabilistic, and that the brain makes use of strong prior beliefs in order to interpret these sensory signals [Mittelstaedt, etc]. A further challenge faced by the vestibular system is that it needs to distinguish between sensory signals that are the result of active motion, and those resulting from external forces [@Cullen:2009dc, @Cullen:2014gx]. The ability to distinguishing re-afferent from ex-afferent sensory signals is an essential capability of all organisms that move actively [@Crapse:2008jc].

Furthermore, the apparent connection between mental imagery and predictive processing has repeatedly been alluded to [@Moulton:2009ij, @Clark:2012vf, @Gambi:2015gv], and @Grush:2004wx claims that mental imagery is performed by emulator circuits implementing a forward model. These ideas can also be addressed from a computational point of view in the context of dynamic Bayesian inference, where there are multiple roles for sensory predictions, over different time scales. This was recently discussed by [@Tian:2012ui] in the context of imagined speech. 
<!-- [ok] -->
The putative link between predictive processing and higher cognitive processing has a firm basis in predictive processing accounts of sensory inference [Friston]. The predictive processing framework has been applied to almost all sensory modalities [CITATIONS: pain, proprioception, auditory , visual, etc.], and the picture that has emerged in recent years is that the brain employs sophisticated inference algorithms in order to "interpret" the stream of incoming sensory data. The brain seems to be actively engaged in formulating hypotheses, and testing these on sensory data; this has been extended to the concept of active inference [Friston], in which an agent attempts to either change it's internal model of world, or to interven on the world in order to fulfil sensory predictions.

We claim that the vestibular modality is particularly well-suited to investigating predictive processing at multiple hierarchical scales, from the lowest levels of sensory inference to higher levels, where goals, intentions and beliefs of the organism become relevant. A case in point is the study by @Wertheim:2001cb, who provided evidence that participants' perceptions of tilt induced by vestibular stimulation are shaped to a large extent by the beliefs about possible motions they formed prior to taking part in the experiement.
<!-- [ok] -->

Interference effects between higher-level cognitive processes, such as SPT and mental imagery, and sensory inference, are thought to occur as a result of shared circuitry [@Deroualle:2014ho, @Nigmatullina:2015gh]. However, the precise nature of such interactions, and what exaclty might be shared, is left unspecified. At first glance, it appear unlikely that such interference should be possible, as low levels of vestibular processing appear to be cognitively impenetrable [CITATION]. In particular, @Nigmatullina:2015gh found that the onset of the vestibulo-ocular reflex (VOR) was earlier when the direction of an imagined self-rotation was congruent with the direction of a subsequently experienced rotation. The VOR, however, must be susceptible to the influence of intentions and goals; for example, if a person intends to fixate a target during a movement, the VOR should be cancelled. Thus, seemingly reflexive, and impenetrable, motor actions require a large amount of flexibility based on top-down information. Even accepting the fact that low level sensorimotor processing may be more cognitively penetrable than previously thought, the fact that imagining a rotation of the self should have an effect on a person's ability to detect rotations, or to discriminiate between directions of rotation, still requires an explanation.

> NOTE: we need to include all studies demonstrating interference, especially in the direction cognition -> inference (Nigmatullina effect)

@Ellis:2017fg recently claimed that the types of interference effects might arise as a result of cognitive processing exerting an influence on circuits performing sensory inference via shared higher-level priors. In particular, the interference described by @Nigmatullina:2015gh looks like an effect of a bias in the sensory discrimination task. Although vestibular perceptual thresholds have been extensively investigate for yaw rotations [@Merfeld:2011fs; @Grabherr:2008kk], the methods used for estimating thresholds require unbiased responding by the participants. There have no investigations of biasing effects on yaw rotations.

In this paper, we pursue the claim that interference should be investigated as a bias, and we attempt to explain the grounding of higher level cognitive abilities in the computations performed by low-level sensory circuits in the service of dynamic real-time interactions with the world. 
<!-- [ok] -->

@Kwisthout:2016fx make the point that predictive processing accounts are situated at low levels in the processing hierarchy, usually at the level of sensory inference or at the level describing kinematics.

In this paper, we address this issue by starting from a model developed for low-level sensory inference, incorporating basic abilities to predict sensory consequences of movement based on knowledge about the kinematics, and extending this to account for interactions between thought processes, operationalized as mental simulations of sensorimotor experience, and sensory processing.

@Battaglia:2013fm addressed the idea that the higher level cognitive ability to perform mental rotation are grounded in the circuits performing sensory inference. We claim that if vestibular cognition is grounded in sensory inference, and that this is to be experimentally studied by analyzing interactions between thought and sensory inference, then it is vital that this be carried within the constraints of a theoretical (computational) framework. A suitable formalism for this is the framework of probabilistic graphical models [@Pearl:1988wz, @Koller:2009ty], and within this framework, interactions between thought and perception can be modelled as effects of higher level priors and hyperpriors on inference at a computational level. Investigating inference at the level of algorithmic implementations might provide further insight.

Probabilistic graphical models are able to represent hierarchically structured knowledge [@Kwisthout:2016fx, @Griffiths:2012ii], which is required for Bayesian models of cognition [@Tenenbaum:2011hd, @Lucas:2010tu]. This approach has been used to model concept learning [@Kemp:2007eo], inductive biases in reasoning, in cognitive models of development [@Perfors:2011ex], language [@Goodman:2016go] and social cognition [@Baker:2009kx].

In this paper, we focus specifically on the role of the vestibular system for spatial perspective transformations, and we draw the link between probabilistic computations performed by the vestibular system, and the brain's ability to run mental simulations for the purpose of adopting the perspective of another person.

@@

Following @Karmali:2012cv, we focus on one specific type of head movement, a rotation of the head about the yaw axis, and discuss this in the context of the probabilistic computations that are performed when the goal is to:

@@

<!-- [ok] -->

@@

## Previous work
Whereas early models of vestibular sensory processing were inspired by control theory [@Borah:1988fd, @Merfeld:1999cg], several recent studies have described vestibular sensory processing as dynamic Bayesian inference [@Laurens:2006be, @PaulRMacNeilage:2008bv, @Karmali:2012cv]. @Selva:2012gsa
provide a recent review of the relationship between optimal observer models and models that employ Kalman filtering, and show that these are largely equivalent. Both @Karmali:2012cv and @Laurens:2006be described vestibular sensory inference in terms of particle filtering, which is a particular algorithm that can be used to perform inference in state-space models. [@Doucet:2000bh, @Doucet:2009us]. Using their model, @Laurens:2006be were able to simulate perceptual responses to centrifugation and off-vertical axis rotations. A well-known phenomenon of vestibular perception is velocity storage, which is defined as the slower decay of perceptual and oculomotor responses, relative to the SCC afferent responses [@Raphan:1979hn], __[Robinson 1977]__. This is generally seen as evidence for the existence of an internal model in the processing of vestibular signals. @Karmali:2012cv focussed on one-dimensional angular velocity during yaw rotations; their model was able to explain the phenomenon of velocity storage as a consequence of particle filtering, with the time constant of velocity storage depending on the noise of the sensory afferent signals.

## Generative model for sensory inference
We proceed by providing a brief introduction to the concept of state-space models, and particle filtering, which is a particular algorithm that can be used to perform inference in state-space models. State-space models can be seen as special cases of more general Bayesian hierarchical models [@Bishop:2006ui; @Murphy:2012ua]. Following the same approach as @Penny:2013iv, we first reframe the generative model used for sensory inference as a dynamic probabilistic graphical model, which is suitable for modelling the time-varying behaviour of a system in terms of latent, unobservable processes and noisy observations.

For example, consider trying to estimate the velocity and position during a leftward head turn. The afferent SCC signals are considered to be proportional to the angular velocity of the head with a frequency band of approximately 0.1 – 10 Hz, which constitutes the range of natural head movements [@Grabherr:2008kk; @Carriot:2014fv]. The measurements are therefore observations of the true, unobservable angular velocity; however, they are contaminted by Gaussian noise, which is determined by characteristics of the sensory organs. Figure \@ref(fig:head-turn-data) shows a sequence of noisy measurements during a two-second turn of the head. The acceleration generating the movement is sinusoidal, given for each time step $t$ by $Asin(2\pi t/T)$, where $A$ is the amplitude of the movement, $T = 1/f$ is the duration of the movement, and $f$ is the frequency. This acceleration results in a peak velocity $\omega_{peak} = AT/\pi$ and a final angular position $\theta_{final} = AT^2/2\pi$.  Particle filtering provides a means of filtering out the noise inherent in the measurements, and allows estimation of evolution of the unobservable, latent state variables.

@@

motiondata_hidden <- motiondata %>%
    gather(type, value = value, -time) %>%
    filter(type != "observations") %>%
    mutate(type = factor(type,
                         levels = c("position",
                                    "velocity",
                                    "acceleration")))

p1 <- motiondata_hidden %>%
        ggplot(aes(x = time, y = value)) +
            geom_line(aes(linetype = type), size = 1.4) +
            geom_point(data = motiondata,
                       aes(x = time, y = observations),
                       shape = 21, colour = "white",
                       fill = "white", size = 6, stroke = 2) +
            geom_point(data = motiondata,
                       aes(x = time, y = observations),
                       shape = 21, colour = "black",
                       fill = "white", size = 4, stroke = 2) +
            guides(linetype = guide_legend(keywidth = 2.5, keyheight = 1)) +
            xlab("Time (sec)") +
            ylab("Angular position (deg)") +
            theme_apa() +
            theme(legend.title=element_blank())
p1
```

@@

How do we explain higher-level anticipation/expectations in such a model? This form of anticipation is not the same as the 1-step ahead predictions made during inference, but rather implement higher level prior beliefs about the direction of movement, for example.

### 2017-01-27 notes.Rmd
draft introduction

interference arises when he brain, due to task demands, uses the control mechanism in a manner which is not entirely counterfactual

# two
interference is surprising, because a mental simulation should be counterfactual. if it isn't this might be a problem. why should thinking about rotating left have an effect on your ability to sense leftward motion?

we show this type of interactions needs to be investigate pdf in a probabilistic framework

@@

in any case, we need a closer look at the connection between menral simulation and planning and the ability to predict self initiated movements.

>>>>

__interference arises when he brain, due to task demands, uses the control mechanism in a manner which is not entirely counterfactual__

# two
~~Interference is surprising, because a mental simulation should be counterfactual. if it isn't this might be a problem. why should thinking about rotating left have an effect on your ability to sense leftward motion?~~

~~we show this type of interactions needs to be investigate pdf in a probabilistic framework~~

@@

in any case, we need a closer look at the connection between mental simulation and planning and the ability to predict self initiated movements.
